Coming up to this point is one of the biggest challenge I have ever pursued
in my life. This thesis would not have been possible if I did not have the 
support of the following people.

I owe my deepest gratitude to my adviser, Prof Vikram Adve, whose guidance, patience and encouragement
have been pivotal in successful completion of this thesis. He is one of the nicest \& smartest
person I have ever met in my life. He has an ocean-deep of patience to listen to all my ideas and
never ceased to amaze me with his clever insights and thoughtful suggestions. 
Also, it is worth mentioning the huge amount of importance that he yields to the wellbeing  of his
students. I could not have imagined a better mentor during this journey. 
Thanks for providing me with this lovely memories! 

I owe special thanks to Theodoros Kasampalis, Daejun Park, Sushant Dinesh, Edward J. Schwartz and Will Dietz
with whom I have collaborated at some point for research. I have enjoyed the many discussions we have had on 
our work and on research life in general.

I owe my thanks to all my committee members and all my colleagues in LLVM research group for
their valuable feedback, ideas and discussions. 

I would like to thank the \K team, for their technical support throughout the project, and
the Strata \& Mcsema developers, for promptly confirming our reported bugs and answering
all our questions in great detail. I am also grateful to Alastair Reid and Matthew Fernandez for 
their invaluable feedback.

Last but not the least, I am grateful to my dearest family - my parents, sister and my wife,
for their love, support and understanding all these years.